Pillar 1 - Sequential Pattern Mining

### 2019-01-28 Pillar 1 - Sequential Pattern Mining.Rmd
solve conflict
creates rules with grades

Now we include data for pass and fail:

```{r apriori PF}

# creating vector of fail course
course_id_fail <- d_transactions_PF %>%
  filter(PF == "fail") %>%
  distinct(`Course ID`)

@@

```{r, cache = TRUE}
tmp_distributed_transcript <- d_transcript %>%
  filter(!is.na(Grade), Grade>=0) %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T)
#This is the transcript with cumulative error gap.
distributed_transcript <- spread(tmp_distributed_transcript, key = "Round_grade", value = Values #, fill = F
               )%>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10,
         ) %>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  select(-Gr_bool )%>%
  separate(Gr_string, into=c("Trash", "Grade_new"), sep="g")%>%
  select(-Trash) %>%
  mutate(Grade= as.numeric(Grade_new))%>% #can put this into Grade and have just one grade value.
  select(-Grade_new)%>%
  mutate(Item=paste(`Course ID`,Grade, sep="_"))%>%
  select(`Student ID`,`Course ID`)%>%
  arrange(`Course ID`, `Student ID`) %>%

```
#puting everything into quasi-function
--------------------------------------------------
```{r}

#parameters for analysis
category <-quo(Low_grade) #Outcome or Low_grade or Discrete_grade or #None
transcript_to_use <- "distributed_transcript" #d_transcript or distributed_transcript 
fail_threshold <- 6.5

#Preparing data
d_sequence_transactions <- d_course %>%

  left_join(get(transcript_to_use), by = "Course ID") %>%
  mutate(
    Outcome = 
           case_when( Grade == -1   ~"Fail", #if we don't do this NG satisfies the condition for pass v
                      Grade  > 5.5 ~"Pass",
                               T~"Fail"),
    Low_grade =
           case_when(Grade == -1 ~"low score", #if we don't do this NG satisfies the condition for pass v
                      Grade > fail_threshold ~"high score",
                               T~"low score"),
    Discrete_grade =
           case_when(Grade <= 5.4             ~ 5.0,
                     between(Grade, 5.5, 6.4) ~ 6.0,
                     between(Grade, 6.5, 7.4) ~ 7.0,
                     between(Grade, 7.5, 8.4) ~ 8.0,
                     between(Grade, 8.5, 9.4) ~ 9.0,
                     Grade >= 9.5             ~ 10.0,
                     T ~ -99999    #this is a control, we should not have any grades here
                     ),
    None= "" # this might not be optimal
         ) %>%

@@

#checking results
as(transactions, "data.frame") #WARNING: THIS TAKES A LOT OF TIME, FOR SOME REASON
```
##Create filter condition.
```{r}
string_filter <- case_when(quo_name(category)=="Outcome"        ~ "Fail",
                           quo_name(category)=="Low_grade"      ~ "low score",
                           quo_name(category)=="Discrete_grade" ~ "5",
                           quo_name(category)=="None"           ~ "_" #useless filter
                           )
```
##CSPADE Algorithm:
```{r}
frequent_trajectories <- cspade(transactions, parameter = list(support = 0,
                                                               maxsize = 1,
                                                               maxlen = 2 #,
                                                               #mingap=, #gap between consecutive elements
                                                               #maxgap=, #hap between consecutive elements
                                                               #maxwin=  #gap between ANY two elements
                                                                 ), control   = list(verbose = TRUE))

@@

  filter(str_detect(RHS, string_filter)) %>% #FILTER (YES) --> uncomment, (NO)-->comment

@@

>>>>

```{r apriori PF}
# creating vector of fail course
course_id_fail <- d_transactions_PF %>%
  filter(PF == "fail") %>%
  distinct(`Course ID`)

@@

```{r}
tmp_distributed_transcript <- d_transcript %>%
  filter(!is.na(Grade), Grade>=0) %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T)
#This is the transcript with cumulative error gap.
distributed_transcript <- spread(tmp_distributed_transcript, key = "Round_grade", value = Values #, fill = F
               )%>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10,
         ) %>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  select(-Gr_bool )%>%
  separate(Gr_string, into=c("Trash", "Grade_new"), sep="g")%>%
  select(-Trash) %>%
  mutate(Grade= as.numeric(Grade_new))%>%
  select(-Grade_new)%>%
  filter(!str_detect(Period, " to "))%>% #Crutch FILTER
  mutate(Item=paste(`Course ID`,Grade, sep="_"))%>%
  mutate( Time = paste(Year_numerical, Period, sep = ""),
          Time = as.double(Time))%>%
  select(`Student ID`, Item, Time) %>%
  arrange(`Student ID`, Time) %>% #Note that integer identifiers must be positive and that transactions must be ordered by sequenceID and eventID.
  group_by(`Student ID`, Time) %>%
  summarize(list_course = list(Item)) %>%
  mutate(TransactionID = paste(`Student ID`, Time, sep = "_"))

@@

```{r}

tmp_distributed_transcript <- d_course %>%
  select(`Course ID`, Type, Letters) %>%
  # Exclude 
  filter(
    Type != "Mandatory",                  # (i) mandatory courses e.g. COR, CAP, etc
    ! Letters %in% c("SKI", "PRO",        # (ii) skills & projects (taken by majority of students)
                     "SAS", "SAH", "SAC") # (iii) courses of semester abroad (uninformative)
    ) %>%

  # Join with transcripts.
  inner_join(d_transcript, by = "Course ID") %>%
  filter(!is.na(Grade), Grade>=0) %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T)
#This is the transcript with cumulative error gap.

distributed_transcript <- spread(tmp_distributed_transcript, key = "Round_grade", value = Values #, fill = F
               )%>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10,
         ) %>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  select(-Gr_bool )%>%
  separate(Gr_string, into=c("Trash", "Grade_new"), sep="g")%>%
  select(-Trash) %>%
  mutate(Grade= as.numeric(Grade_new))%>%
  select(-Grade_new)%>%
  filter(!str_detect(Period, " to "))%>% #Crutch FILTER
  mutate(Item=paste(`Course ID`,Grade, sep="_"))%>%
  mutate( Time = paste(Year_numerical, Period, sep = ""),
          Time = as.double(Time))%>%
  select(`Student ID`, Item, Time) %>%
  arrange(`Student ID`, Time) %>% #Note that integer identifiers must be positive and that transactions must be ordered by sequenceID and eventID.
  group_by(`Student ID`, Time) %>%
  summarize(list_course = list(Item)) %>%
  mutate(TransactionID = paste(`Student ID`, Time, sep = "_"))

#puting everything into quasi-function
--------------------------------------------------
##to transaction
```{r}
transaction_list_course <- distributed_transcript$list_course
names(transaction_list_course) <- distributed_transcript$TransactionID

transactions <- as(transaction_list_course, "transactions")
transactionInfo(transactions)$sequenceID <- distributed_transcript$`Student ID`
transactionInfo(transactions)$eventID <- distributed_transcript$Time
transactionInfo(transactions)$transactionID <- NULL

#checking results
as(transactions, "data.frame")
```
##Create filter condition.
```{r}
# string_filter <- case_when(quo_name(category)=="Outcome"        ~ "Fail",
#                            quo_name(category)=="Low_grade"      ~ "low score",
#                            quo_name(category)=="Discrete_grade" ~ "5",
#                            quo_name(category)=="None"           ~ "_" #useless filter
#                            )
```
##CSPADE Algorithm:

```{r}
frequent_trajectories <- cspade(transactions, parameter = list(support = 0.1,
                                                               maxsize = 1,
                                                               maxlen = 2 #,
                                                               #mingap=, #gap between consecutive elements
                                                               #maxgap=, #hap between consecutive elements
                                                               #maxwin=  #gap between ANY two elements
                                                                 ), control   = list(verbose = TRUE))

@@

  #filter(str_detect(RHS, string_filter)) %>% #FILTER (YES) --> uncomment, (NO)-->comment

@@

support(frequent_trajectories,transactions, control=list(verbose=T))
